---
title: "Data Tidying Project"
author: "Kevin Gilds"
date: "6/13/2018"
output: html_document
---


This project was completed in coordination with Chromebook Data Science Course Data Tidying. 



The chunk below loads required packages and installs them if not already done so.
```{r setup, include=FALSE}
## install packages if necessary
list.of.packages <- c("httr", "readxl", "dplyr", "tidyr", "stringr", "reshape2", "skimr", "ggplot2")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

## load packages
library(httr)
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(reshape2)
library(skimr)
library(ggplot2)
```

### Get Data

[Data Set 1](https://data.world/retail/department-store-sales): Sales from the Retail Trade and Food Services Report from the US Census. This dataset only covers Department Stores, though the report covers a wide range of retail types. [1992-2016]

[Data Set 2](https://data.world/garyhoov/retail-sales-growth) US Retail Sales by Store Type with Growth Rate [2009-2014]


This chunk uses the GET function to obtain the data from the link and reads the excel file with the read_excel function. 
```{r data, warning=FALSE} 
#1992-2016
#https://data.world/retail/department-store-sales

GET("https://query.data.world/s/gdk7iwtlisq6vkktmybqqr7hjjty5s", write_disk(tf <- tempfile(fileext = ".xls")))
df1 <- read_excel(tf)

#2009-2014
# https://data.world/garyhoov/retail-sales-growth
GET("https://query.data.world/s/py7kinxvyuxjpzwdjs2ti4wdmui6bi", write_disk(tf <- tempfile(fileext = ".xls")))
df2 <- read_excel(tf)
## the the first row and make that the column names of the data frame
colnames(df2) <- df2[1,]
```

### Save Raw Data

This code chunk save the data to the repository in the even the data is moved from its location on the internet. Notice that it was required to use a relative path to get from the code directory to the raw-data directory

```{r save-data, warning=FALSE} 
## use saveRDS() to save each object as a .rds file 



saveRDS(df1,"../../data/raw_data/df_department.rds")
saveRDS(df2,"../../data/raw_data/df_retail.rds")
```

### Wrangle Data

This code chunk was pre-written. It cleans up column names. If you notice some variables names are year with an r at the end. This chunk also filters data pertaining to Retail sales, total and department stores. 
```{r wrangle-df2, warning=FALSE} 
## an example working with df2
## let's wrangle!
df_retail <- df2 %>%
  ## remove the r from the column names of df2
  magrittr::set_colnames(gsub("r","",df2[1,])) %>% 
  ## add a new column called "business"
  mutate(business = gsub("[…]|[.]","",`Kind of business`)) %>%
  ## filter to include Retail sales or Department stores sales
  filter(grepl('Retail sales, total |Department stores', business)) %>%
  ## only look at columns with year information in them
  select(.,c(matches('19|20'),business)) %>%
  ## take year column and collapse them into a single column
  gather(., "year", "n", 1:(ncol(.)-1)) %>%
  ## make sure the count column `n` is numeric
  mutate(n=as.numeric(n)) %>%
  ## filter to only include the businesses we're interested in
  filter(business == "Retail sales, total "| business=="Department stores ")
```

The code chunk below works through df2 by separating the period variable into a more tidy format. This code chunk works to make variable names consistent in order to merge with df2. 
```{r wrangle-df1, warning=FALSE} 
## now, your turn!
## work with df1
df_department <- df1 %>% 
  ## split Period column into one column called "month" and one called "year"
  separate(Period, into = c("month", "year")) %>%
  ## add a column `value` which contains the 
  ## information from the `Value (in millions)` 
  mutate(value = `Value (in millions)`) %>%
  ## group the data frame by the `year` column
  group_by(year) %>%
  ## Summarize the data by creating a new column
  ## call this column `n` 
  ## have it contain the sum of the `value` column
  summarise(n = sum(value)) %>% 
  ### create a new column called `business`
  ## set the value of this column to  be "department stores" 
  ## for the entire data set 
  mutate(business = "department stores") %>%
  ## reorder column names to be : business, year, n
  select(business, year, n)
```

This code chunk uses the rbind function to merge the two datasets. 
```{r merge-data, warning=FALSE} 
## Now, combine the two data frames
df_total <- rbind(df_department, df_retail)
```

This plot was already created via the course.
```{r plot, warning=FALSE, fig.width=10} 
## Plot Retail Sales data
ggplot(df_retail, aes(x=year,y=n,colour=business)) +
  geom_point() 

## Plot Department Sales data
ggplot(df_department, aes(x=year,y=n)) +
  geom_point() 

## Plot Combined Data
ggplot(df_total, aes(x=year,y=as.numeric(n), colour=business)) +
  geom_point()
```